---
title: 'Lab 7: Data Visualization and Collaborative Coding'
author: "DOAgoons"
date: "2023-03-02"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lab Overview

This lab is all about data visualization. Data Viz is a powerful tool for informing and persuading audiences and you will also learn how to craft interactive data visualizations, which are appropriate for presentations as well as web-based dashboards.

## Technical Learning Objectives

- Understand the basic structure of the grammar of graphics used in ggplot2.
- Understand how to use various Geoms to represent data points and their aesthetic properties to represent variables.
- Understand how to use faceting to divide a plot into subplots based on the values of one or more discrete variables.
- Understand how ggplot2 integrates with the tidyverse and RPubs, RMarkdown, and RShiny.

## Business Learning Objectives
- Understand the role of data visualization (data viz) in data science and analytics.
- Understand multiple strategies and options for presenting data through effective visualizations (data viz).
- Understand the power of faceting to illustrate comparisons amongst variables
- Understand the continued advances RStudio is making in data viz with ggplot and htmlwidgets for interactive data viz.

There are three main parts to the lab:

- Part 1: Data Visualization Using ggplot2, guides you through an understanding of the basic structure of the grammar of graphics, including the use of various Geoms to represent data points and their aesthetic properties to represent variables. It will also help you to understand how to use faceting to divide a “plot“ into “subplots” based on the values of one or more discrete variables. You will also learn how to better use the coordinate systems, as well as building and customizing labels and legends.

- Part 2 (Optional): Plotting Data Visualizations on Maps, focuses on data visualizations on maps. Like with the social media lab earlier, this part of the lab is not required. *If you want to complete this section, you will need to acquire your own Google Maps API key from Google*. Given our uses, this API should be available at no charge from Google (but you will be required to add a method of payment).

- Part 3: Interactive Data Visualizations, introduces you to htmlwidgets. This package htmlwidgets was developed by RStudio to allow you to build rich, interactive, data visualizations. These data visualizations are suitable for publishing on the web with RPubs, RMarkdown, or RShiny.


&&&& download dataset left 

## Installing Packages, Importing and Exploring Data

For this lab you will be working primarily in ggplot2

Deliverable 1: Get your working directory and paste below: Deliverable 2: Load the Required libraries
```{r}

getwd()
### Install Packages 

#install.packages("geojsonio")
#install.packages("highcharter")
#install.packages("leaflet")
#install.packages("plotly")
# install.packages("Rserv")
#install.packages("sf")
#install.packages("ggthemes")
#install.packages("DT")
#install.packages("gapminder")
#install.packages("devtools")
#install.packages("remotes")

### Load packages 


library(highcharter)
library(leaflet)
library(plotly)
library(remotes)
library(Rserve) # this package installed compared to Rserv which didn't
library(devtools)
library(sf)
library(ggthemes)
library(DT)
library(gapminder)
library(geojsonio)
library(tm)
library(tidytext)
library(ggmap)
library(ggthemes)

# When installing, you may be asked if you want to update all related packages. Generally, the recommendation is to select the number 1 to update all the related packages.

### Now using the remotes package, install oidnChaRts. You can use this line of code if you wish: 
remotes::install_github("martinjhnhadley/oidnChaRts")
library(oidnChaRts)


# Comment: This package allows you to visualize how a specific htmlwidget library will plot/display your data. It also comes with several built in datasets for plotting.

# Install the new version of ggmao from git hub
```

```{r}
# Once you install the devtools package, load the devtools library and then install the most recent version of gmaps from github. You can use this line of code: 

devtools::install_github("dkahle/ggmap", force = TRUE)
```


For Part 02 of the lab, you need to get a Google API. 

To begin, visit: http://console.developers.google.com/apis/dashboard (http://console.developers.google.com/apis/dashboard)
You must log into your Google account if you are not already logged in; then accept the terms (if you wish) and click the create project button. NB: This is where you will need to enable a billing plan for the project, or you will receive error messages later. This may take a little while, and involve new steps if Google changes the API. You next create and enable the Google APIs we will use. We need an API key, so click on that link, and you may search within the API library. There are many APIs from which to choose; we want; (1) Static Maps and (2) Geocode).
Now click on create credentials and choose the API key option (you will see there is also an Oath option, which you may need for other packages. Save that information somewhere safe, and remember not to share that key with anyone (including me).


This part of the lab will guide us through an understanding of the basic structure of the grammar of graphics. RStudio’s ggplot2 is part of the tidyverse and loads when the tidyverse is loaded. In ggplot2, you use this grammar of graphics to build every graph from the same key components, which are: 
- A dataset; 
- A coordinate system; and 
- Geoms
With these three components, you may customize any data visualization using ggplot2. This section of the lab also helps you to understand how to use “faceting” to divide a “plot“ into “subplots” based on the values of one or more discrete variables. You will also learn how to better use the coordinate systems, as well as building and customizing labels and legends.

The ggplot2 package allows you to build and customize data visualizations; tweak those visualizations to make them look exactly as you would like it. This “Grammar of Graphics” allows us to construct sentences for data visualization that others can understand.

Within the Grammar of Graphics there are four components necessary for you to understand how to build compelling visualizations. Each of those components plays a different role.
Component (argument) Description/Purpose
1. Data (data) = Raw data to visualize
2. Geometries (geom) = shapes to visualize data
3. Aesthetics (aes) = properties of geometries (size, color, etc) 4. Scale (scale) = mapping between geo and aes

Before you begin data visualization, you must first have the data structured in a way you want to be able to analyze it. To do so, you may need to manipulate and filter your data (for example, with dplyr) to create a data frame consisting of only the data you want to use for your visualization.

Deliverable 1: Get your working directory and paste below:
```{r}
getwd()
```
Deliverable 2: Load the Required libraries


Deliverable 3: Get and Explore the Diamonds Dataset
```{r}
data("diamonds")
head(diamonds)
tail(diamonds)
```
Deliverable 4: Create a Histogram of “diamonds” in Base R
```{r}
hist(diamonds$carat, main = "Carat Histogram", xlab = "Carat")
```

Deliverable 5: Ccreate a Scatterplot of “diamonds” using Base R
```{r}
plot(price~carat, data=diamonds)
```
Deliverable 7: Create a Blank Canvas in ggplot
```{r}
ggplot()
```
Deliverable 8: Rebuild the diamonds histogram with ggplot
```{r}
ggplot(data=diamonds) + geom_histogram(aes(x=carat))
```
Deliverable 9: Build a Density Plot
```{r}
ggplot(data=diamonds) + geom_histogram(aes(x=carat), fill = "grey50")
```
Deliverable 10: Build a Scatterplot with ggplot
```{r}
ggplot(diamonds,aes(x=carat, y=price)) + geom_point()
```
Deliverable 11: Re-Build Your Scatterplot with Color
```{r}
# This prints a whiteboard
g = ggplot(diamonds,aes(x=carat, y=price))

# Now add points to the printed whiteboard. This prints a scatterplot in black & white colors
g + geom_point() 

# Now, we make the above visual more appealing to decision makers by adding colors 
g + geom_point(aes(color=color))
```
Notice that the colored scatter is plotted for 7 categories of a variable. WE can visualize each of these categories separately by using #facet_wrap(~color)

Deliverable 12: Demonstrate a Facet Wrap Plot
```{r}
# Now, we make the above visual more appealing to decision makers by adding colors 
g + geom_point(aes(color=color)) + 
  facet_wrap(~color)
```
Deliverable 13: Demonstrate a Facet Wrap Grid Plot
```{r}
g + geom_point(aes(color=color)) + 
  facet_grid(~color)
```
```{r}
g + geom_point(aes(color=color)) + 
  facet_grid(cut~color)
```

Deliverable 14: Demonste a Facet Wrap Plot with a Histogram
```{r}
ggplot(diamonds, aes(x = carat)) + geom_histogram() 

ggplot(diamonds, aes(x = carat)) + 
    geom_histogram() + 
  facet_wrap(~ color)

```
Deliverable 15: Demonstrate Boxplots in ggplot
```{r}

# Take a look at the variables in the Diamonds dataset
variable.names(diamonds)

summary(diamonds)

# Plots blanck slate
ggplot(diamonds, aes(y=carat, x=1)) 


# Now, add a boxplot to the slate
ggplot(diamonds, aes(y=carat, x=1)) +
  geom_boxplot()


# Take a look at diamonds$cut vs diamonds$carat in general
ggplot(diamonds, aes(y=carat, x=diamonds$cut)) +
  geom_boxplot()

# Take a look at diamonds$cut vs diamonds$carat for all the categories 
ggplot(diamonds, aes(y=carat, x=diamonds$cut)) +
  geom_boxplot() +
  facet_wrap(~color)
```

Deliverable 16: Demonstrate a Violinplot in ggplot
```{r}
ggplot(diamonds, aes(y=carat, x=cut)) +
  geom_violin()
```
Deliverable 17: Read in a CSV file from the Web, Explore, and Summarize
```{r}
college = read.csv("http://672258.youcanlearnit.net/college.csv")
```
Exploratory analysis: We will look at the imported data head, tail and do a high level summary of the data. We will later on check for missing values and verify if there are any duplicated rows in the imported dataset. Finally, we will use the create_report() from the DataExplorer library which summarizes the above exploratory analysis visually. 
```{r}


# Look at the head and tail of imported data
head(college)
tail(college)
summary(college)
class(college)
# is.na(college)
# duplicated(college)
attach(college)

library(DataExplorer)
# create_report(college)
```
Observation: the variables name, city, state, region, highest_degree, control, gender, loan_default_rate need to be converted to type, 'categorical' to get a better print out of the summary of the data. Basically, based on your general culture, you'll know which variables to convert to which data type. 

Deliverable 18: Wrangle the Data and Re-Summarize
```{r}
college <- college %>% 
  mutate(name=as.factor(name), city=as.factor(city), state=as.factor(state), region=as.factor(region), highest_degree=as.factor(highest_degree),control=as.factor(control), gender=as.factor(gender),loan_default_rate=as.numeric(loan_default_rate))

summary(college)
# create_report(college)
```
You may see a warning, “NA’s introduced by coercion”. This means, when there is missing data, it introduces Base R’s standard NA to represent missing data.

Print the variables of this dataset and do prior analysis for univariate and bi-variate analysis. ANOVA and regression will also be good techniques to analyse this data
```{r}
names(college)
```

Deliverable 19: Create a Basic Scatterplot of the college Data
```{r}

# Lets do the scatterplot step by step
ggplot(data=college)

# Populate slate
ggplot(data=college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg))
```
# Deliverable 20: Add Shape Control to Your Scatterplot. This is a black/white visual which is not very appealing. 
```{r}
ggplot(data=college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg, shape=control))
```

Deliverable 21: Add Color to Your Shape Control. *Replace shape of the scatter to color*. Rem, the points on the scatter were all uniform until we used different variables in the dataset to have a better understanding of the data. 
```{r}
ggplot(data=college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg, color=control))

# We can separate the scatter on two separate graphs using facet_wrap
ggplot(data=college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg, color=control)) +
  facet_wrap(~control)

ggplot(data=college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg, color=control)) +
  facet_grid(~highest_degree)

ggplot(data=college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg, color=highest_degree))
head(college)

ggplot(data=college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg, color=highest_degree)) +
  facet_grid(control~highest_degree)

ggplot(data=college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg, color=highest_degree)) +
  facet_grid(highest_degree~control)
```

Deliverable 22: Add Size to Your AES and Interpret. We add size to undergrads
```{r}
ggplot(data=college) + 
  geom_point (mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads))
```
Deliverable 23: Add Alpha (Opacity) to Your AES

Now, we can make the distinction between these points clearer by adding alpha to the geom call. Adding alpha, sets the percentage of opacity of the dots. Adding the alpha outside the first geom call, means the alpha is not being set based on any level/values in the data.
alpha=1 (100% opaque, 0% transparent)
alpha=1 / 100 (1% opaque, 99% transparent)
Keep modifying the alpha value until you get the result you want.

As an example you may use the following code: ggplot(data = college) + geom_point(mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads), alpha=1/2)
```{r}
ggplot(data = college) + 
  geom_point(mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads), alpha=1/2)
```

Deliverable 24: Add a Line Plot and Smooth 
```{r}
ggplot(data = college) + 
  geom_line(mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads), 
             alpha=1/2)
```

```{r}

ggplot(data = college) + 
  geom_smooth(mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads), 
             alpha=1/2)

ggplot(data = college) + 
  geom_boxplot(mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads), 
             alpha=1/2)
```
 You can add multiple geom (plot types) on the same plot (eg lines and plots)


Deliverable 25: Create a Bar Graph

Now, let’s create a bar graph: 
- geom_bar() = creates a bar graph, specify the x value; count is the y value

- geom_col() = creates a column graph
user can set both the x and y value
That is the only difference.
To add values on the bar/column, we add fill=control (to fill in the bar by that variable)
As an example you may use the code below:
that is, we are plotting a bar graphs of region and then seeing who in that region is a subset of control variable (public or private). NOTE THAT, This code will use default colors when filling the bars in the bar graph. You can always change them as you please. 
```{r}
# ggplot(data = college) +  geom_bar(mapping = aes(x=region, fill=control)) 

ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) + 
  theme(panel.background = element_blank()) + theme(plot.background = element_blank()) + scale_x_discrete(name="Region") + scale_y_continuous(name = "Number of Schools", limits = c(0,500))
```

Deliverable 26: Change Colors on your Bar Graph

To change the colors on our bar graph, add a new indented line with a "+" and then use the *scale_fill_manual()* function with the argument containing a concatenated list of colors.
*As an example, you may use this line of code:* scale_fill_manual(values = c(“orange”,“blue”))
```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) + 
  theme(panel.background = element_blank()) + 
  theme(plot.background = element_blank()) + 
  scale_x_discrete(name="Region") + 
  scale_y_continuous(name = "Number of Schools", limits = c(0,500)) + 
  scale_fill_manual(values = c("grey","blue"))
```

Deliverable 27: Add Legends and Titles to Your Bar Graphs

*To adjust the legends on your graph*, use guide=guide_legend(title ="Institution Type")). We change the legend title from "Control", which is the name given to the variable with levels *Private, Public* to "Institution Type" which speaks more to the end user. simply add the argumnent to the code after a comma "," sign.  
```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) + 
  theme(panel.background = element_blank()) + 
  theme(plot.background = element_blank()) + 
  scale_x_discrete(name="Region") + 
  scale_y_continuous(name = "Number of Schools", limits = c(0,500)) + 
  scale_fill_manual(values = c("grey","blue"), guide = guide_legend(title ="Institution Type"))

```
You may also adjusting the legends on the graphs by making the legend on 1 row. To do this, add nrow=1 at the end of the argument.

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) + 
  theme(panel.background = element_blank()) + 
  theme(plot.background = element_blank()) + 
  scale_x_discrete(name="Region") + 
  scale_y_continuous(name = "Number of Schools", limits = c(0,500)) + 
  scale_fill_manual(values = c("grey","blue"), guide = guide_legend(title ="Institution Type", nrow = 1))
```
You may also change the label position. Add one more element to the end of the argument after nrow=1. use label.position = “bottom”.

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) + 
  theme(panel.background = element_blank()) + 
  theme(plot.background = element_blank()) + 
  scale_x_discrete(name="Region") + 
  scale_y_continuous(name = "Number of Schools", limits = c(0,500)) + 
  scale_fill_manual(values = c("grey","blue"), guide = guide_legend(title ="Institution Type", nrow = 1, abel.position = "bottom"))
```

You may continue to modify the legend on the graph by adjusting size of legend box. To do so, add one more element to the end of the arugment after label.position. This element should be keywidth = 2.5. This makes the keywidth broader and freater. 
```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) + 
  theme(panel.background = element_blank()) + 
  theme(plot.background = element_blank()) + 
  scale_x_discrete(name="Region") + 
  scale_y_continuous(name = "Number of Schools", limits = c(0,500)) + 
  scale_fill_manual(values = c("grey","blue"), guide = guide_legend(title ="Institution Type", nrow = 1, abel.position = "bottom", keywidth = 2.5))
```

*You may adjust the location of the legend.* After the previous last line, add a + and on the next indented line call the theme() function, with legend.position. = “top” in the argument. In your code, pay close attention to where the plus signs + are located at the end of the lines.
```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) + 
  theme(panel.background = element_blank()) + 
  theme(plot.background = element_blank()) + 
  scale_x_discrete(name="Region") + 
  scale_y_continuous(name = "Number of Schools", limits = c(0,500)) + 
  scale_fill_manual(values = c("grey","blue"), guide = guide_legend(title ="Institution Type", nrow = 1, abel.position = "bottom", keywidth = 2.5)) +
  theme(legend.position = "top")
```

Deliverable 28: Add Annotations to Your Bar Graphs

Annotations for Data Visualizations 
- annotate() adds text and shape 
- geom_hline() adds horizontal line 
- geom_vline() adds vertical lines 
- Place a text label “Elite Privates” on the graph
As an example, you may use the following code:

```{r}
# ggplot(data = college) + 
 # geom_point(mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads), alpha=1/2) + 
 # annotate("text", label="Elite Privates", x=45000, y=1450) + 
 # geom_hline(yintercept = mean(college$sat_avg)) + 
 #  annotate("text", label ="MeanSAT ", x = 47500, y = mean(college$sat_avg)-15) +
 #  geom_vline(xintercept = mean(college$tuition)) + 
 #  annotate("text", label = "MeanTuition", y = 700, x = mean(college$tuition) +7500)
```

You may make the chart more visually appealing by removing the panel background. As an example you may use the following code:

```{r}
 # ggplot(data = college) + 
  # geom_point(mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads), alpha=1/2) + 
 #  annotate("text", label = "Elite Privates", x = 45000, y = 1450) + 
  # geom_hline(yintercept = mean(college$sat_avg)) + 
  # annotate("text", label ="MeanSAT ", x = 47500, y = mean(college$sat_avg)-15) +
  # geom_vline(xintercept = mean(college$tuition)) + 
 #  annotate("text", label = "MeanTuition", y = 700, x = mean(college$tuition) +7500) + 
 #  theme(panel.background = element_blank(), legend.key = element_blank()) 
```

To continue making the chart more visually appealing you may change the size. As an example you may use the following code:
```{r}
 # ggplot(data = college) + 
  # geom_point(mapping=aes(x=tuition, y=sat_avg, color=control, size=undergrads), alpha=1/2) + 
  # annotate("text", label = "Elite Privates", x = 45000, y = 1450) + 
  # geom_hline(yintercept = mean(college$sat_avg)) + 
 #  annotate("text", label ="MeanSAT ", x = 47500, y = mean(college$sat_avg)-15) +
 #  geom_vline(xintercept = mean(college$tuition)) + 
 #  annotate("text", label = "MeanTuition", y = 700, x = mean(college$tuition) +7500) + 
 #  theme(panel.background = element_blank(), legend.key = element_blank()) + 
 #  scale_color_discrete(name="Institution Type") + 
  # scale_size_continuous(name="Undergraduate") + 
  # scale_x_continuous(name = "Tuition") + 
 #  scale_y_continuous(name="SAT Score")
```

For additional adjustments on your graph, you may the location of the legend. As an example, you may use the following code: (THIS CODE IS OKAY)

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) + 
  theme(panel.background = element_blank()) + 
  theme(plot.background = element_blank()) + 
  scale_x_discrete(name="Region") + 
  scale_y_continuous(name = "Number of Schools", limits = c(0,500)) + scale_fill_manual(values = c("orange", "blue"),guide=guide_legend(title="Institution Type", nrow=1, label.position = "bottom", keywidth = 2.5)) + 
  theme(legend.position = "bottom") + ggtitle("More college are in the Southern USA than any other region", subtitle = "Source: U.S. Department of Education")
```

Deliverable 29: Explore Pre-Defined Themes for Your Bar Graphs
Using Pre-Defined Themes
There are some pre-defined themes in R, but there are libraries available that contain many more. Let’s start with our bar graph, but rewind to when we first created it
Example:
```{r}
ggplot(data = college) + geom_bar(mapping = aes(x=region, fill=control))
```
This presents the ggplot default theme. Now, we will switch to a theme called bw. To do so, add a plus + at the end of the line, and on the next indented line add the theme_bw() function. 

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_bw()
```

For this and each of the next several codes, substitute the theme_bw() function for the indicated theme based function.
theme_minimal()
```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_minimal() 
```

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_void()
```


```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_dark()
```

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_solarized()
```

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_solarized()
```

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_excel()
```


```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_wsj()
```

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_economist()
```

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_economist_white()
```

```{r}
ggplot(data = college) + 
  geom_bar(mapping = aes(x=region, fill=control)) +
  theme_fivethirtyeight()
```


### Part 3: Interactive Data Visualizations

Finally, in Part 3 of the lab, we will introduce you to htmlwidgets. This package htmlwidgets was developed by RStudio to allow you to build rich, interactive, data visualizations. These data visualizations are suitable for publishing on the web with RPubs, RMarkdown, or RShiny.

The htmlwidgets are loaded separate packages, which are: highcharter (makes interactive charts), leaflet (makes interactive maps), plotly (makes Java Script visualizations, DT (makes interactive tables), and sf.
```{r}
# install.packages("highcharter") 
# install.packages("leaflet")
# installed.packages("plotly")
# install.packages("DT")

library(highcharter) # Makes Interactive Charts
library(leaflet) # Makes Interactive Maps
library(plotly) # Makes Java Script Visualizations
library(DT) # Makes Interactive Tables
library(sf) # Makes simple features 
```
You must separately load each htmlwidget library before using it. However, here is a package called oidnChaRts that allows you to compare an interactive barchart in highcharter and one of the other htmlwidgets library. It also comes with several built-in datasets for plotting.


CREATE AND VIEW THE DATA WE WILL PLOT
To begin, create some data we can use to try out these mapping options. You may use the example code below:

```{r}
data_to_plot <- data_stacked_bar_chart %>% 
  group_by(country_group, occupation) %>% 
  summarise(total = sum(count)) %>% 
  ungroup()
```

Deliverable 30: View Data in Highcharter
Let’s start by viewing our data_to_plot object in the highcharter package. You must separately load each htmlwidget library before using it. You may use the sample code below:

```{r}
library(highcharter) 

data_to_plot %>% 
  stacked_bar_chart(library = "highcharter", 
                    categories.column = ~ country_group, 
                    subcategories.column = ~ occupation, 
                    value.column = ~ total )
```
Deliverable 31: View Data in Plotly

Now try the plotly package. You may use the code below: library(plotly)

```{r}
data_to_plot %>% 
  stacked_bar_chart(library = "plotly", 
                    categories.column = ~ country_group, subcategories.column = ~ occupation, value.column = ~ total) %>% 
  layout(margin = list(l = 150), yaxis = list(title = ""))
```

Deliverable 32: View Scatter Geomarkers
Now try Scatter Geo Markers.
```{r}
data_geo_marker_plot %>% 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(label = ~city)
```

Deliverable 33: View Scatter geoplots
```{r}
data_geo_marker_plot %>% 
  leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(radius = ~count/8, label = ~city)
```

Deliverable 34: View Clustered Scatter Geoplots
Now try Clustered Scatter Geo Plots.

```{r}
data_geo_marker_plot %>% 
  leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(radius = ~count/8, label = ~city, clusterOptions = markerClusterOptions())
```
Deliverable 35: Read in Shapefiles and View in Cloropleth Map
You may also read in shapefiles, in both esri and json formats. 

```{r}
getwd()

esri_shapefile <- read_sf("world-shape-files")

geojson_shapefile <- read_sf("world-geojson.json")

shp_as_sf <- read_sf("world-shape-files")

converted_geojson <- geojson_list(shp_as_sf)

gapminder_most_recent <- gapminder %>% 
  filter(year == max(year)) 
highchart(type = "map") %>% 
  hc_add_series_map(map = converted_geojson, df = gapminder_most_recent, value = "pop", joinBy = c("name_long", "country")) %>% 
  hc_title(text = "Population Choropleth for 2007 (Source: Gapminder)")
```








